---
title: "PS 7"
author: "Luca Moretti"
date: "17 Mai 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 7.1


```{r cars, message=FALSE, warning=FALSE}
library(foreign)
library(readstata13)
md <- read.dta13("runandjump_sample1500g.dta")
md$bwtcent = md$bweight-1500
```

b)

```{r b, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(data=md)+
  geom_histogram(aes(bwtcent),binwidth = 1, center=.5)
```

c)
You have selection into treatment.

d)
.. sorry I ran out of time, the exam is so close Ahhhhhhhhhhhh


## 7.2 Applications
As we understand the paper they want to analyze the question: whether institutions shape culter or vice versa. They want to see if an institutional change (the policy) leads to a convergence in culture (measured as time between children).



a)
Treatment: cultural background * policy implementation
We believe they want to estimate the heterogenous effect of the policy implementation on various groups of immigrants vs locals. We think that in this case culture is a collectionterm for groupfixed effects. Therefore, all observed and unobserved differences enter.

b)
The types that emmigrate may be different than the ones that stay in their country.

c)
They were thinking to difference them out and estimate a different effect of the policy on the group. The problem is that the group dummy interacted with the treatment brings back all the unobserved and observed variables. 
We have to assume that the all other characteristics in which they differ interact with the policy. Only culture creates variation in the reaction to the policy change.

d)
If it were a proper diff-in-diff this would make me more confident that the trends after the intervention would have been parallel as well. 
However, with regard to the assumption stated in c) this doesn't help us too much. Hence internal validity needs extremely stron assumption that in my opinion are not given e.g. if a migrant group has a lower income (probably this is the case), then lower income families cannot react stronger to the incentive set by the policy.

e)
Maybe you could control for observable interactions. But you would still have the unobservables problem mentioned in their abstract.






```{r pressure, echo=FALSE}
```

